---
title: "Google scholar info for IAE scholars"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Working with RStudio and R has developed over the past few years to include many of the tools needed to create reproducible and reuseable documents, packages and functions. I hope this repository will add to these example for researchers at the University of Canberra.

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## Data

For this project the data includes the peer reviews publications generated by the Institute of Applied Ecology and includes the code needed to update and investigate these publications using the following packages:

```{r}

```

## Including Code

You can include R code in the document as follows:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
